---
title: 'Reproducible Research: Course Project 1'
author: "Yih-Kuang Lu"
date: "11/13/2020"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

It is now possible to collect a large amount of data about personal movement using activity monitoring devices such as a Fitbit, Nike Fuelband, or Jawbone Up. These type of devices are part of the “quantified self” movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. But these data remain under-utilized both because the raw data are hard to obtain and there is a lack of statistical methods and software for processing and interpreting the data.

This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.

The data for this assignment can be downloaded from the course web site:

Dataset: Activity monitoring data [52K]
The variables included in this dataset are:

steps: Number of steps taking in a 5-minute interval (missing values are coded as \color{red}{\verb|NA|}NA)
date: The date on which the measurement was taken in YYYY-MM-DD format
interval: Identifier for the 5-minute interval in which measurement was taken
The dataset is stored in a comma-separated-value (CSV) file and there are a total of 17,568 observations in this dataset.

```{r}
# download data from source and unzip data
file<-'repdata_data_activity.zip'
download.file('http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip',file,method='curl')
unzip(file)

# Read data and take a look of data
data <- read.csv("activity.csv")
head(data)
```

## What is mean total number of steps taken per day?
1.    Calculate the total number of steps taken per day
2.    If you do not understand the difference between a histogram and a barplot, research the difference between them. Make a histogram of the total number of steps taken each day
3.    Calculate and report the mean and median of the total number of steps taken per day
```{r}
# get the sum of the steps w/ respect to date variable
steps_perday <- aggregate(steps ~ date, data, sum)
head(steps_perday)
```

```{r}
# generate a histogram for steps per day
hist(steps_perday$steps, main = paste("Steps per Day"),xlab="Number of Steps")
```

```{r}
# average (or mean) of Steps per day
rmean <- mean(steps_perday$steps,na.rm=TRUE)
rmean
```

```{r}
# median of Steps per day
rmedian <- median(steps_perday$steps,na.rm=TRUE)
rmedian
```

## What is the average daily activity pattern?
1.    Make a time series plot (i.e. \color{red}{\verb|type = "l"|}type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
2.    Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
# get the average of steps per date w/ respect to interval variable
steps_interval <- aggregate(steps ~ interval, data, mean)
head(steps_interval)
```

```{r}
# generate a plot for averaged Steps per date w/ interval
plot(steps_interval$interval,steps_interval$steps, type="l", col='red', xlab="Interval", ylab="Number of Steps",main="Averaged Steps per Day w/ Interval")
```

```{r}
# get the interval which has the max averaged Steps
max_interval <- steps_interval[which.max(steps_interval$steps),1]
max_interval
```

## Imputing missing values
Note that there are a number of days/intervals where there are missing values (coded as \color{red}{\verb|NA|}NA). The presence of missing days may introduce bias into some calculations or summaries of the data.

1.    Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with \color{red}{\verb|NA|}NAs)
2.    Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.
3.    Create a new dataset that is equal to the original dataset but with the missing data filled in.
4.    Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?
```{r}
# calculate tooal missing data in the dataset
total_missing <- sum(!complete.cases(data))
total_missing
```

```{r}
# get the average of steps w/ date 
ave_steps <- aggregate(steps ~ interval, data, mean)
new_addmissing <- NULL
for (i in 1:nrow(data)) {
    tmp <- data[i, ]
    if (is.na(tmp$steps)) {
        steps <- subset(ave_steps, interval == tmp$interval)$steps
    } else {
        steps <- tmp$steps
    }
    new_addmissing <- c(new_addmissing, steps)
}
head(new_addmissing)
```

```{r}
# update (or create) new dataset by adding NA step data w/ date variable
new_data<-data
new_data$steps <- new_addmissing
head(new_data)
```

```{r}
# generate a histogram for steps per day using new data
new_steps_perday <- aggregate(steps ~ date, new_data, sum, na.rm = TRUE)
hist(new_steps_perday$steps, main = paste("Steps per Day"), col='red', xlab="Number of Steps")
# overlap w/ the histogram using original data 
hist(steps_perday$steps, main = paste("Steps Each Day"), xlab="Number of Steps", add=T)
legend("topright", c("Imputed", "Non-imputed"), col=c("red", "grey"),pch=c(20,20),cex=1)
```

```{r}
# average (or mean) of Steps per day using new data
rmean <- mean(new_steps_perday$steps,na.rm=TRUE)
rmean
```

```{r}
# median of Steps per day using new data
rmedian <- median(new_steps_perday$steps,na.rm=TRUE)
rmedian
```

```{r}
# generate difference in the average of Steps per day between original and new data
rmediandiff <- mean(new_steps_perday$steps,na.rm=TRUE) - mean(steps_perday$steps,na.rm=TRUE)
rmediandiff
```

```{r}
# generate difference in the median of Steps per day between original and new data
rmeandiff <- median(new_steps_perday$steps,na.rm=TRUE) - median(steps_perday$steps,na.rm=TRUE)
rmeandiff
```

## Are there differences in activity patterns between weekdays and weekends?
For this part the \color{red}{\verb|weekdays()|}weekdays() function may be of some help here. Use the dataset with the filled-in missing values for this part.

1.    Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
2.    Make a panel plot containing a time series plot (i.e. \color{red}{\verb|type = "l"|}type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.

```{r}
# generate the steps data by sorting date variable into weekday and weekend
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
new_data$week = as.factor(ifelse(is.element(weekdays(as.Date(new_data$date)),weekdays), "Weekday", "Weekend"))
steps_week <- aggregate(steps ~ interval + week, new_data, mean)
head(steps_week)
```

```{r}
library(ggplot2)
library(ggpubr)
p1<-ggplot(steps_week[which(steps_week$week=='Weekday'),] , aes(x = interval , y = steps)) + geom_line(col='red') + ggtitle("The average of Steps in weekdays") + scale_x_continuous("Interval") + scale_y_continuous("Number of Steps")
p2<-ggplot(steps_week[which(steps_week$week=='Weekend'),] , aes(x = interval , y = steps)) + geom_line(col='blue') + ggtitle("The average of Steps in weekends") + scale_x_continuous("Interval") + scale_y_continuous("Number of Steps")
ggarrange(p1,p2,nrow=2)

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
